---
title: "Monte Carlo Simulation"
author: "Lucia Bonfanti"
format: html
editor: 
  markdown: 
    wrap: sentence
---

(Adapted from Monte Carlo Simulation: An Introduction for Engineers and Scientists by Alan Stevens, Chapter 11)

Polymers are large molecules comprised of a large number of repeated units (monomers) linked end to end.
There are naturally occurring polymers, such as cellulose, and artificial polymers, such as polythene (polyethylene).
A simple model of a polymer assumes its linked units are each oriented completely randomly with respect to each other.
This results in a nearly infinite number of configurations.

Suppose the goal is to estimate the straight-line distance between the start and end of the polymer chain.
The theoretical probability density function for this distance is $$P(R) = 4\pi R^2\left(\frac{3}{2\pi\langle r^2\rangle}\right)^{3/2}\exp\left(-\frac{3R^2}{2\langle r^2\rangle}\right), $$ where $R$ is the end-to-end distance, and the mean square position of the units is $\langle r^2\rangle = N u^2$ where $N$ is the number of units in the chain and $u$ is ithe length of each unit.
For simplicity, we'll work with $u=1$.

# Unconstrained Polymer Chains

The position of each link of the chain relative to the previous link can be described with angles $\theta\in[0,2\pi]$ and $\phi\in[0,pi]$ (think of $\theta$ as describing the direction in the XY plane and $\phi$ as describing the direction in the plane uniquely defined by the points $(0,0,0)$, $(x, y, 0)$, and $(x, y, z)$).
In this case, $\phi$ need only cover $\pi$ radians.

Without loss of generality, we can assume that the chain starts at the origin.
Then, the end point in the chain can be described by a cumulative sum of a series of x, y, and z coordinates that are a function of $\theta_{n\times 1}$ and $\phi_{n\times 1}$:

$$\begin{array}{rcl} x &=& \cos(\theta)\sin(\phi)\\
y &=& \sin(\theta)\sin(\phi)\\
z &=& \cos(\phi)\end{array} $$

The distance between the origin and the end of the chain $(x,y,z)$ can be calculated as $\sqrt{x^2 + y^2 + z^2}$.

## Single Unconstrained Link Function

Write a function that takes $n$ and returns the distance between the start and end points of a chain.
Use vector-based operations to increase the efficiency of your function.

## Distribution of Unconstrained Links

Simulate 10000 chains and save the distribution of the end-to-end distance.

```{r db-setup-r}
functionchain <- function(n) {
  theta <- runif(n, min=0, max=2*pi)
  phi<- runif(n,min=0,max=pi)
  x <- cos(theta)*sin(phi)
  y <- sin(theta) * sin(phi)
  z <- cos(phi)
  X=sum(x)
  Y=sum(y)
  Z=sum(z)
  result <- sqrt(X^2+Y^2+Z^2)
  return(result)
}

set.seed(999)
functionchain(100)

chain <- data.frame (x = replicate(10000, functionchain(100)))
library(ggplot2)
ggplot(data = chain, aes(x=x)) + geom_histogram()

```



# Constrained Polymer Chains

For real polymers, adjacent links don’t have the complete freedom of orientation allowed in the case of the ideal model.
We can simulate a simple constrained orientation model by restricting the relative orientation of adjacent links.
We’ll prevent adjacent links from orienting themselves such that there is an acute angle smaller than, say 45° between them.
We can do this by checking the angle between adjacent links after choosing a random orientation and repeatedly choosing the random orientation until that angle is larger than 45°.

## Single Constrained Link Function

Write a function that takes $n$ and returns the distance between the start and end points of a chain.
Can you think of a way to use vectorized functions to accomplish this task?
Why or why not?
No, because the link will depend on the previous one 
## Distribution of Constrained Links

Simulate 10000 chains and save the distribution of the end-to-end distance.
Plot the density of the unconstrained links and the density of constrained links on the same plot -- how much has the distribution changed?
 It change, but not so much. 

```{r db-setup-r}
distance_con <- function(n, min_angle = pi/4) {

  theta <- runif(1, min=0, max=2*pi)
  phi <- runif(1, min=0, max=pi)
  chain <- matrix(c(cos(theta)*sin(phi), sin(theta)*sin(phi),cos(phi)), nrow=1)
  
 
  for (i in 2:n) {
    repeat {
      theta <- runif(1, min=0, max=2*pi)
      phi <- runif(1, min=0, max=pi)
      new_link <- c(cos(theta)*sin(phi),sin(theta)*sin(phi),cos(phi))
      
      prev_link <- chain[i-1, ]
      angle <- acos(sum(prev_link * new_link) / (sqrt(sum(prev_link^2)) * sqrt(sum(new_link^2))))
      
      if (angle >= min_angle) break
    }
    chain <- rbind(chain, new_link)
  }
  result <- sqrt(sum(colSums(chain)))

}
set.seed(999)
chainc <- data.frame (x = replicate(10000, distance_con(100,min_angle = pi/4)))
library(ggplot2)
ggplot(chainc, aes(x=x)) + geom_histogram()

```

